---
title: "Hands-on Exercise 8"
author: "Sei Sar Hla Kyi"
date: "16 October 2024"
date-modified: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# **14  Geographically Weighted Predictive Models**

## **Installing and Loading R packages**

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

## Importing and Preparing Data

```{r}
mdata <- read_rds("data/mdata.rds")
```

## Data Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

```{r}
#|eval: FALSE
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
#|eval: FALSE
write_rds(train_data, "data/train_data.rds")
write_rds(test_data, "data/test_data.rds")
```

## **Computing Correlation Matrix**

Examine if there is sign of multicolinearity

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

::: callout-note
## The correlation matrix above shows that all the correlation values are below 0.8. Hence, there is no sign of multicollinearity.
:::

## **Building a non-spatial multiple linear regression**

```{r}
train_data <- read_rds("data/train_data.rds")
test_data <- read_rds("data/test_data.rds")
```

```{r}
#|eval: FALSE
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
#|eval: FALSE
write_rds(price_mlr, "data/price_mlr.rds" ) 
```

## **GWR predictive method**

### **Converting the sf data.frame to SpatialPointDataFrame**

```{r}
train_data_sp <- as_Spatial(train_data)
train_data_sp
```

### **Computing adaptive bandwidth**

Next, `bw.gwr()` of **GWmodel** package will be used to determine the optimal bandwidth to be used.

```{r}
#|eval: FALSE
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data_sp,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

The result shows that 40 neighbour points will be the optimal bandwidth.

### **Constructing the adaptive bandwidth gwr model**

```{r}
#|eval: FALSE
write_rds(bw_adaptive, "data/bw_adaptive.rds")
```

```{r}
bw_adaptive <- read_rds("data/bw_adaptive.rds")
```

Calibrate the GWR-based hedonic pricing model by using adaptive bandwidth and Gaussian kernel as shown in the code chunk below.

```{r}
#|eval: FALSE
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data_sp,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```

```{r}
#|eval: FALSE
write_rds(gwr_adaptive, "data/gwr_adaptive.rds")
```

### **Retrieve GWR output object**

```{r}
gwr_adaptive <- read_rds("data/gwr_adaptive.rds")
gwr_adaptive
```

### **Converting the test data from sf data.frame to SpatialPointDataFrame**

```{r}
test_data_sp <- test_data %>%
  as_Spatial()
test_data_sp
```

### **Computing adaptive bandwidth for the test data**

```{r}
gwr_bw_test_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data_sp,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

### **Computing predicted values of the test data**

```{r}
set.seed(123)
train_data_sample <- train_data_sp[sample(1:nrow(train_data_sp), 1000), ]
test_data_sample <- test_data_sp[sample(1:nrow(test_data_sp), 500), ]
```

```{r}
gwr_model <- gwr.predict(
    formula = resale_price ~ floor_area_sqm + storey_order + remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN + WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH, 
    data = train_data_sample, 
    predictdata = test_data_sample, 
    bw = 40, 
    kernel = "gaussian", 
    adaptive = TRUE, 
    longlat = FALSE
)
```

## **Preparing coordinates data**

### **Extracting coordinates data**

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
coords_train <- write_rds(coords_train, "data/coords_train.rds" )
coords_test <- write_rds(coords_test, "data/coords_test.rds" )
```

### **Droping geometry field**

We will drop geometry column of the sf data.frame by using `st_drop_geometry()` of sf package.

```{r}
train_data <- train_data %>% 
  st_drop_geometry()
```

## **Calibrating Random Forest Model**

```{r}
#|eval: FALSE
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data)
rf
```

```{r}
#|eval: FALSE
write_rds(rf, "data/rf.rds")
```

```{r}
rf <- read_rds("data/rf.rds")
rf
```

## **Calibrating Geographical Random Forest Model**

### **Calibrating using training data**

The code chunk below calibrate a geographic ranform forest model by using `grf()` of **SpatialML** package.

```{r}
#|eval: FALSE
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                       remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

```{r}
#|eval: FALSE
write_rds(gwRF_adaptive, "data/gwRF_adaptive.rds")
```

```{r}
gwRF_adaptive <- read_rds("data/gwRF_adaptive.rds")
```

### **Predicting by using test data**

```{r}
test_data <- cbind(test_data, coords_test) %>%
  st_drop_geometry()
```

Next, `predict.grf()` of spatialML package will be used to predict the resale value by using the test data and gwRF_adaptive model calibrated earlier.

```{r}
#|eval: FALSE
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

```{r}
#|eval: FALSE
GRF_pred <- write_rds(gwRF_pred, "data/GRF_pred.rds")
```

Converting the predicting output into a data frame

```{r}
GRF_pred <- read_rds("data/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

```{r}
test_data_p <- cbind(test_data, GRF_pred_df)
```

```{r}
write_rds(test_data_p, "data/test_data_p.rds")
```

### **Calculating Root Mean Square Error**

The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis.

```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

### **Visualising the predicted values**

```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```

::: callout-note
## A better predictive model should have the scatter point close to the diagonal line. The scatter plot can be also used to detect if any outliers in the model.
:::
