---
title: "Take Home Exercise 3: Modelling Geography of Financial Inclusion in Tanzania with Geographically Weighted Methods "
author: "Sei Sar Hla Kyi"
date: "14 October 2024"
date-modified: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# 1 Overview

## 1.1 Background

Financial inclusion plays a key role in promoting equitable economic growth by enabling individuals and businesses to access essential financial services such as banking, loans, insurance, and equity products. In Tanzania, financial inclusion initiatives have become a priority in addressing the needs of underserved populations, particularly those in rural or economically disadvantaged areas. These initiatives aim to improve household incomes, reduce income inequality, and foster inclusive economic participation by providing timely and affordable access to financial products and services(United Nations Capital Development Fund \[UNCDF\], 2024).

Despite significant progress in financial inclusion efforts, access to financial services is uneven across different geographical regions within Tanzania. Rural areas, particularly in districts with low levels of economic activity, often face greater challenges in accessing financial services compared to more urbanized districts. These spatial disparities raise important questions about the factors driving financial inclusion at the district level and how regional policies can be designed to promote more inclusive economic outcomes.

## 1.2 Objectives

This analysis aims to build an explanatory model using Geographically Weighted Regression (GWR) to identify the key factors affecting financial inclusion across districts in Tanzania. By focusing on multiple socio-economic and infrastructural variables, this study will provide insights into how these factors vary geographically and influence financial access at a district level.

Specifically, the objectives are:

-   **Identify Key Determinants**: Determine the primary socio-economic and infrastructural factors that influence financial inclusion, such as income levels, education, proximity to financial institutions, and mobile money adoption.

-   **Examine Regional Variation**: Explore how the effects of these factors vary across different districts, providing a localized understanding of financial inclusion.

# 2 Methodology

The explanatory model will be developed using Geographically Weighted Regression (GWR), which is a spatial extension of multiple linear regression that allows for localized variations in the relationships between predictors and the response variable—in this case, financial inclusion.

The process will follow these steps:

-   **Defining dependent variable (y)**: Derive dependent variable which represents financial inclusion from the FinScope dataset.

-   **Explanatory Variable Selection (x)**: Identify key predictor variables from the FinScope dataset, such as income levels, education and mobile money adoption. These will serve as explanatory variables in the model.

-   **Model Building**: Build the explanatory model using multiple linear regression techniques, enhanced by GWR to account for geographical variations. This will allow us to evaluate the influence of each predictor on financial inclusion across different districts.

-   **Model Calibration and Validation**: Calibrate the GWR model to ensure that it effectively captures the localized relationships between financial inclusion and its determinants. The model will be validated by examining the accuracy and goodness-of-fit metrics.

-   **Spatial Visualization**: Mapping the results of the GWR model to visualize how the effects of each factor vary geographically, will highlight districts where specific variables have a stronger or weaker influence on financial inclusion.

------------------------------------------------------------------------

# 3 Loading R Packages

::: panel-tabset
## Packages Used

The following packages were used for this study:

+-------------+---------------------------------------------------------------------------------------------------------------+
| Package     | Description                                                                                                   |
+=============+===============================================================================================================+
| olsrr       | For creating and evaluating ordinary least squares regression models.                                         |
+-------------+---------------------------------------------------------------------------------------------------------------+
| ggstatsplot | for producing statistical visualizations that combine ggplot2 graphics with statistical results               |
+-------------+---------------------------------------------------------------------------------------------------------------+
| ggpubr      | To create publication-ready visualizations with ggplot2 and arranging multiple plots                          |
+-------------+---------------------------------------------------------------------------------------------------------------+
| sf          | For importing, managing, and handling geospatial data                                                         |
+-------------+---------------------------------------------------------------------------------------------------------------+
| sfdep       | To analyze spatial relationships and incorporate them into regression models                                  |
+-------------+---------------------------------------------------------------------------------------------------------------+
| spdep       | For calculating spatial weights and conducting tests for spatial relationships                                |
+-------------+---------------------------------------------------------------------------------------------------------------+
| GWmodel     | To explore spatially varying relationships between the dependent and independent variables                    |
+-------------+---------------------------------------------------------------------------------------------------------------+
| tmap        | For thematic mapping, allowing us to visualize spatial data in an intuitive and aesthetically pleasing manner |
+-------------+---------------------------------------------------------------------------------------------------------------+
| tidyverse   | For data manipulation, cleaning, and visualization                                                            |
+-------------+---------------------------------------------------------------------------------------------------------------+
| gtsummary   | To create summary tables for regression results, providing a clear and concise presentation of model outputs  |
+-------------+---------------------------------------------------------------------------------------------------------------+
| see         | For visualizing model diagnostics and results                                                                 |
+-------------+---------------------------------------------------------------------------------------------------------------+
| performance | For evaluating model performance and diagnostics                                                              |
+-------------+---------------------------------------------------------------------------------------------------------------+
| datawizard  | To handle missing values, recode variables, and perform exploratory data analysis before modeling             |
+-------------+---------------------------------------------------------------------------------------------------------------+

## Code

```{r}
pacman::p_load(olsrr, ggstatsplot, ggpubr, sf, spdep, sfdep, GWmodel, tmap, tidyverse, gtsummary,see,performance,datawizard)
```
:::

------------------------------------------------------------------------

# 4 Data Preparation

## 4.1 The Datasets

For this analysis, we will use two primary datasets: (1) FinScope Tanzania 2023, which provides district-level data on financial inclusion, and (2) district-level boundary GIS data from the geoBoundaries portal to facilitate spatial analysis.

-   **FinScope Tanzania 2023**: This dataset include comprehensive district-level data on access to financial services, covering indicators such as banking services, mobile money usage, savings, loans, and insurance. The dataset will be used to identify the key factors affecting financial inclusion and to develop the explanatory model.

-   **District Boundary GIS Data**: Geographic boundary data for the districts of Tanzania, sourced from the geoBoundaries, will enable us to link the financial inclusion data to specific geographic locations for spatial analysis using GWR.

## 4.2 Import the Data

### 4.2.1 Aspatial Data

The *FinScope* *Tanzania* dataset from 2023, was imported by using *read_csv()* package.

Then, the data was prepared for analysis with the following steps:

-   Remove columns with all missing/NA values.
-   Then we select the columns with the variables chosen which will be explained in the Variable Selection section.

```{r}
tanzania <- read_csv("data/rawdata/FinScope_Tanzania.csv")%>%
  select(5,9,11,13:14,16,61,65:67,71,95:96,98:103,107:119,124:125,140,142:151,160:174,177:178,188,183:192,196,215:223,
         275:278,294:306,344:353,361:372,374:378,395:397,455:460,495,497,507:529,574:588,591, 612,650:651,653,655,657,678,680:685,697:698,700:716,718) %>%
  select(where(~ !all(is.na(.))))
```

We will use ***glimpse()*** to quickly check the attributes.

Lastly, we will save the prepared dataframe into rds format.

::: panel-tabset
## Save

```{r}
#| eval: False
write_rds(tanzania, "data/rds/tanzania.rds")
```

## Load

```{r}
tanzania <- read_rds("data/rds/tanzania.rds")
```
:::

### 4.2.2 Geospatial Data - Study Area

The *geo Boundaries of Tanzania* dataset, was imported by using *st_read()* package.

The dataset was prepared for analysis with the following steps:

-   We note that Tanzania is in WGS 84/ UTM Zone 37S. We will use st_transform to reproject coordinate reference system to EPSG: 32737 for Tanzania.

-   To perform left join with the aspatial dataset, we need a common column - districts. Therefore, we need to change the variable, shapeName, to 'dist_name' to make it a common variable with the FinScope Tanzania dataset.

-   To ensure the districts are common, some districts will be renamed from "Arusha Urban" to "Arusha" as the same way it was displayed in FinScope Tanzania dataset. Since the FinScope Tanzania dataset already have a variable that classifies if the surveyee is residing in an urban or rural area, we do not need to have a separate district with the title Urban as it falls under the same district.

```{r}
tanzbounds <- st_read(dsn = "data/rawdata/", 
                      layer = "geoBoundaries-TZA-ADM2",
                      crs=4326) %>%
  st_transform(crs=32737) %>%
  rename(dist_name = shapeName) %>%
  mutate(dist_name = ifelse(dist_name == "Arusha Urban", "Arusha", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Meru", "Arumeru", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Iringa Urban","Iringa", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Bukoba Urban", "Bukoba", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Mpanda Urban", "Mpanda", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Kigoma  Urban", "Kigoma", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Moshi Urban", "Moshi", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Lindi Urban", "Lindi", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Musoma Urban", "Musoma", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Mbeya Urban", "Mbeya", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Babati UrbanBabati Urban", "Babati", dist_name)) %>%
  mutate(dist_name = ifelse(dist_name == "Dodoma Urban", "Dodoma", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Kibaha Urban", "Kibaha", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Morogoro Urban", "Morogoro", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Mtwara Urban", "Mtwara", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Njombe Urban", "Njombe ", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Mtwara Urban", "Mtwara", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Shinyanga Urban", "Shinyanga", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Singida Urban", "Singida", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Songea Urban", "Songea", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Sumbawanga Urban", "Sumbawanga", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Tabora Urban", "Tabora", dist_name))%>%
  mutate(dist_name = ifelse(dist_name == "Tanga Urban", "Tanga", dist_name))
```

```{r}
st_crs(tanzbounds)
```

::: panel-tabset
## Save

```{r}
#| eval: False
write_rds(tanzbounds, "data/rds/tanzbounds.rds")
```

## Load

```{r}
tanzbounds<- read_rds("data/rds/tanzbounds.rds")
```
:::

### 4.2.3 Left Join

Next, we will create a tourism economy indicators layer by left joining the geospatial dataframe (tanzbound) to the attribute dataframe (tanzania) by using ***left_join()***.

::: callout-note
## To perform a left_join(), we must ensure that there is a common column between the two. For our case, the common column is 'dist_name'
:::

```{r}
#| eval: False
tanzania_fin <- left_join(tanzbounds,tanzania)
```

# 5 Deriving the Dependent Variable

## **5.1 Financial Inclusion in Tanzania**

Financial inclusion refers to the process of ensuring that individuals and businesses have access to affordable and useful financial products and services that meet their needs—ranging from transactions, payments, savings, and credit to insurance. According to the World Bank, financial inclusion is a "key enabler in reducing poverty and boosting shared prosperity by providing the unbanked and underserved populations access to essential financial services" (World Bank, 2023).

A robust financial inclusion system allows individuals to manage risks, invest in education, and build sustainable livelihoods. In Tanzania, financial inclusion is especially critical for rural populations, where access to traditional banking infrastructure is limited. Many rely on informal financial services, which can hinder their ability to seize economic opportunities. The growth of digital financial services, particularly mobile money, has dramatically expanded financial inclusion across Sub-Saharan Africa, enabling millions to access financial products without needing to visit a physical bank (World Bank Group, 2024).

Despite significant progress, the uneven distribution of financial services between rural and urban areas underscores the need to examine the factors driving financial inclusion at a local level. Understanding these determinants is essential for enhancing access to financial services and fostering inclusive economic development. This study aims to derive a dependent variable representing financial inclusion from the FinScope dataset, focusing on key financial behaviors.

## **5.2 The Dependent Variable - Financial Inclusion Index**

Research by Demirgüç-Kunt and Klapper (World Bank,2018) indicates that financial inclusion is multi-dimensional, encompassing access to credit, savings, and various financial services. Composite indices, such as the Global Financial Inclusion Index, often aggregate multiple financial behaviors to capture the complexity of financial inclusion. For this study, we will construct a composite measure of the **access to credit and** **financial support** to represent financial inclusion in Tanzania.

The dependent variable will be derived from **total borrowings** and **government/NGO support**, for the following reasons:

1.  **Borrowing Amounts**:

    Borrowing is a crucial indicator of financial inclusion as it reflects an individual's ability to access formal financial institutions. Borrowing, particularly from formal sources like banks or cooperatives, allows individuals to invest in business ventures, manage emergencies, and smooth consumption. Access to credit is widely regarded as a critical dimension of financial inclusion, as it integrates individuals into the financial system. Studies by World Bank (2018) and other financial inclusion research confirm that borrowing from formal institutions directly correlates with increased financial inclusion. In Tanzania, access to loans is an important means by which individuals participate in the financial system, particularly in rural and low-income areas where formal banking is less accessible.

    -   Borrowing-related Variables:
        -   How much total money did you borrow the last time you borrowed (from Financial Institutions, Employers, MFI, Pension Funds)? in TZS
        -   How much did you borrow the last time you borrowed from the SACCO? TZS
        -   How much did you borrow the last time you borrowed from mobile money? TZS
        -   How much did you borrow the last time you borrowed from a group TZS
        -   How much did you borrow the last time you borrowed from a moneylender? TZS

2.  **Government/NGO Support**:

    Government and NGO support is another essential mechanism that facilitates financial inclusion, particularly for vulnerable populations. Cash transfers and other forms of assistance often require recipients to engage with formal financial services, such as opening bank accounts or using mobile money platforms. This engagement with the formal financial system is a critical step toward broader financial inclusion. Receiving support from government or NGOs not only provides immediate financial relief but also encourages the use of formal financial products and services. Empirical research shows that these programs significantly boost financial inclusion, especially in developing countries, by offering individuals a pathway to financial services that they might otherwise lack access to.

    -   Variable:
        -   The last time you received money (from Government or NGO), how much money did you receive? in TZS

3.  Amount Owed

    Including the amount owed is critical to understanding the financial health of individuals within the context of their borrowing and support. It highlights the individual's debt burden relative to their financial resources. The amount owed is indicative of financial obligations that must be managed, making it a relevant factor in assessing overall financial inclusion.

By focusing on total borrowings, government/NGO support, and the amount owed as core components of the dependent variable, we capture three key dimensions of financial inclusion: access to credit, participation in social safety nets, and the management of debt. These factors are particularly relevant in Tanzania, where many individuals rely on government assistance and microfinance to meet their financial needs. This composite approach aligns with global best practices in financial inclusion research, ensuring a comprehensive measure of overall financial activity.

**Financial Inclusion Index (FII)**, our dependent variable, will be derived with the following formula:

$FII_i =AmountOwed_I/ (TotalBorrowing_i + GovNGOSupport_i)$

where:

-   $AmountOwed_i$ = Total money owed or debt

-   $TotalBorrowing_i$ = Total borrowing amount for individual i

    -   It is calculated by summing up all the borrowing-related variables for individual i

        $TotalBorrowing_i = Borrowings_i + SACCOBorrow_i + MobileMoneyBorrow_i +$

        $GroupBorrow_I + MoneylenderBorrow_i$

        Where:

        -   $Borrowings_i$​ = Amount borrowed from Financial Institutions, MFI, Pension Funds

        -   $SACCOBorrow_i$ = Borrowed from SACCO

        -   $MobileMoneyBorrow_i$​ = Borrowed from Mobile Money

        -   $GroupBorrow_I$ = Borrowed from Social groups

        -   $MoneylenderBorrow_i$ = Borrowed from moneylenders

-   $GovNGOSupport_i​$ = Total money received from government or NGOs by individual i

-   $w_1​,w2​​$ = Weights to balance the contribution of each component

[**Financial Inclusion Index Implications**]{.underline}

This formula allows us to assess an individual's ability to manage their debt in relation to their borrowing capacity and external financial support. A higher value of FII indicates that the amount owed is significant compared to available financial resources, which could imply financial stress or vulnerability. A result of **1 or less** indicates that an individual’s debt is manageable relative to their financial resources, while a result **greater than 1** could indicate a concerning level of debt with less financial resources available. A larger index score will indicate less accessibility to financial products.

A FII score of 0 would indicates a complete lack of financial inclusion for the particular individual. It could be one of the two possibilities:

1.  The individual does not have access to any formal financial services, such as banking accounts, credit, or savings options. This can occur in situations where people are entirely reliant on informal financial systems, like borrowing from friends or family.

2.  The individual does not need to or has never borrowed money or received any form of financial assistance from these sources.

Before we further prepare our variables, we will convert the column names into a more R-friendly format using `make.names()`.

```{r}
#| eval: False
colnames(tanzania_fin) <- make.names(colnames(tanzania_fin))
```

The following code chunk will be used to rename the columns and replace empty spaces of these columns with 0 for further use.

```{r}
#| eval: False
tanzania_fin <- tanzania_fin %>%
  mutate(
    `G.2.6..Approximately.how.much.money.do.you.owe..in.TZS` = 
      gsub("Can't remember|Refused", "0", 
           `G.2.6..Approximately.how.much.money.do.you.owe..in.TZS`, 
           ignore.case = TRUE), 
    AmountOwed = as.numeric(`G.2.6..Approximately.how.much.money.do.you.owe..in.TZS`) 
  ) %>%
  rename(
    GovtNGOSupport = `X3.The.last.time.you.received.money..how.much.money.did.you.receive..in.TZS`,
    Borrowings = `G.2.3..How.much.money.did.you.borrow.the.last.time.you.borrowed..in.TZS`,
    SACCO_Borrow = `X5.2.How.much.did.you.borrow.the.last.time.you.borrowed.from.the.SACCO..TZS`,
    Mobile_Borrow = `X13.How.much.did.you.borrow.the.last.time.you.borrowed..TZS`,
    Group_Borrow = `X4.10.How.much.did.you.borrow.the.last.time.you.borrowed.from.a.group.TZS`,
    Moneylender_Borrow = `X4.How.much.did.you.borrow.the.last.time.you.borrowed.from.a.moneylender..TZS`
  ) %>%
  mutate(
    GovtNGOSupport = ifelse(is.na(GovtNGOSupport), 0, GovtNGOSupport),
    AmountOwed = ifelse(is.na(AmountOwed), 0, AmountOwed),
    Borrowings = ifelse(is.na(Borrowings), 0, Borrowings),
    SACCO_Borrow = ifelse(is.na(SACCO_Borrow), 0, SACCO_Borrow),
    Mobile_Borrow = ifelse(is.na(Mobile_Borrow), 0, Mobile_Borrow),
    Group_Borrow = ifelse(is.na(Group_Borrow), 0, Group_Borrow),
    Moneylender_Borrow = ifelse(is.na(Moneylender_Borrow), 0, Moneylender_Borrow)
  )

```

To ensure all these columns are numeric, we use the following code chunk.

```{r}
#| eval: False
tanzania_fin <- tanzania_fin %>%
  mutate(
    GovtNGOSupport = as.numeric(GovtNGOSupport),
    AmountOwed = as.numeric(AmountOwed),
    Borrowings = as.numeric(Borrowings),
    SACCO_Borrow = as.numeric(SACCO_Borrow),
    Mobile_Borrow = as.numeric(Mobile_Borrow),
    Group_Borrow = as.numeric(Group_Borrow),
    Moneylender_Borrow = as.numeric(Moneylender_Borrow)
  )

```

Finally, we will derive our dependent variable and create a new column **Financial Inclusion Index (FII**). To ensure that division by zero does not occur, this code will also handle cases where `Borrowings` and `GovtNGOSupport` both equal zero.

```{r}
#| eval: False
tanzania_fin <- tanzania_fin %>%
  mutate(
    FII = ifelse((Borrowings+ SACCO_Borrow + Mobile_Borrow+ Group_Borrow + Moneylender_Borrow+ GovtNGOSupport) == 0, 0, AmountOwed / (Borrowings+ SACCO_Borrow + Mobile_Borrow+ Group_Borrow + Moneylender_Borrow+ GovtNGOSupport))
  )
```

::: panel-tabset
## Save

```{r}
#| eval: False 
write_rds(tanzania_fin, "data/rds/tanzania_fin.rds")
```

## Load

```{r}
tanzania_fin<- read_rds("data/rds/tanzania_fin.rds")
```
:::

We will use the following code to view the top 5 FII observations from our newly created dependent variable.

```{r}
top_5_fii <- tanzania_fin %>%
  arrange(desc(FII)) %>%
  select(dist_name, FII) %>%
  head(5)

print(top_5_fii)
```

We see that the top observation is an outlier with a FII value of 2000 while the following observations are around 400 and 300. We will exclude this one observation from our study to ensure that interpretation of our results is not affected.

```{r}
tanzania_fin <- tanzania_fin %>%
  filter(FII != 2000)
```

# **6 Explanatory Variables of Interest**

In the development of a robust multiple linear regression model, the selection of appropriate variables is critical to ensuring the model's accuracy and interpretability. This section focuses on the identification and selection of both nominal and continuous variables that will serve as predictors in our regression analysis.

Initially, we will establish a preliminary set of variables based on theoretical foundations and previous empirical research. Nominal variables will encompass categorical factors that may influence financial behaviors, such as the type of financial institution, demographic characteristics, and access to financial services. Continuous variables will include measurable factors like amount borrowed, and age which can provide insight into the financial landscape of the respondents.

Once we have constructed our initial model using these variables, we will conduct a comprehensive model assessment to evaluate its performance. This will involve calculating the Variance Inflation Factor (VIF) to assess multicollinearity among the independent variables. High VIF values indicate potential redundancy among variables, which can compromise the model's reliability. Consequently, we will refine our variable selection by identifying and retaining only those variables that significantly contribute to explaining the variance in our dependent variable, financial inclusion.

## 6.1 Continuous Variables

The following are the continuous variables selected.

+----------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Variable**                                 | **Justification for Selection**                                                                                                                                                                                                                                                                                                                                                                                 |
+==============================================+=================================================================================================================================================================================================================================================================================================================================================================================================================+
| Age                                          | Age is a critical demographic variable that often influences various socio-economic outcomes. It can affect financial behaviors, access to resources, and the types of assistance individuals might require.Research indicates that younger individuals may be more inclined to take risks and invest in education or business ventures, while older individuals might focus on savings and retirement planning |
+----------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Average Money Spent on Betting               | Studies show that excessive gambling can lead to financial distress and lower overall financial health (Koomson et al., 2022). Lower spending may lead to better financial health and inclusion.                                                                                                                                                                                                                |
+----------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Number of Saving Groups respondent belong to | Higher participation in savings groups typically correlates with better savings rates and financial stability.Groups provide social support and financial education.                                                                                                                                                                                                                                            |
+----------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Contribution to Saving Groups                | Regular contributions can lead to increased savings and financial stability, improving access to credit and financial services.                                                                                                                                                                                                                                                                                 |
+----------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

For these four continuous variables, we will prepare the data in the following steps:

-   Rename the variables

-   Replace NA values with 0

-   Ensure they are in numeric format

-   We found that there are two records that stated that they are in 16,000 and 2000 social groups which is not possible. Therefore, they will be excluded.

```{r}
tanzania_fin <- tanzania_fin %>%
  mutate(
    Group_Belong = as.numeric(replace_na(as.numeric(as.character(X4.3.How.many.savings.groups.do.you.belong.to.in.total.)), 0)),
    Group_Contribution = as.numeric(replace_na(as.numeric(as.character(X4.8.How.much.do.you.contribute.in.total..to.all.the.groups.you.belong..in.a.mont)), 0)),
    AmountBetting = as.numeric(replace_na(as.numeric(as.character(X17c.On.average..how.much.do.you.spend.on.betting.in.a.month.)), 0))
  )%>%
  filter(Group_Belong != 16000)%>%
  filter(Group_Belong != 2000)
```

## 6.2 Nominal Variables

We will employ one-hot encoding to transform nominal variables into a format suitable for regression modeling. One-hot encoding allows us to create binary columns for each category, enabling the regression model to assess the impact of each nominal variable effectively. This approach ensures that we capture the distinct effects of categorical factors on the dependent variable while maintaining the integrity of the data.

We will use the following variables and code chunk to one-hot encode all the nominal variables as mentioned above.

::: panel-tabset
## Variables and Justification

+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Variable**                                                                                                                                                         | **Justification for Selection**                                                    | **Modification**                                                                                                                                                                                         |
+======================================================================================================================================================================+====================================================================================+==========================================================================================================================================================================================================+
| dist_name                                                                                                                                                            | To see the impact of residing in different districts                               | All districts are encoded to different variables starting with `dist_`. For instance, if a respondent is from Babati, the variable `dist_Babati` will be 1 for that respondent and 0 for all the others. |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Cluster type                                                                                                                                                         | To understand if there are any inequality across urban and rural areas             | Encoded as two columns - 'Urban' and 'Rural'                                                                                                                                                             |
|                                                                                                                                                                      |                                                                                    |                                                                                                                                                                                                          |
|                                                                                                                                                                      |                                                                                    | If respondent indicates 'urban', this variable will be 1 and 0 for rural.                                                                                                                                |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Who is the head of this household                                                                                                                                    | To understand the demographic background of respondent                             | Encoded as two columns - 'HHH' for head of household and 'NotHHH' for not head of household                                                                                                              |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Respondent gender                                                                                                                                                    | To understand the demographic background of respondent                             | Encoded as two columns - 'male' and 'female'                                                                                                                                                             |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Education Background                                                                                                                                                 | To understand the demographic background of respondent                             | Encoded as eight columns - 'NoFormalEducation' and 'SomePrimary' and so on                                                                                                                               |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Access to Mobile Phone, Internet and Laptop                                                                                                                          | To understand the socioeconomic status of respondent                               | Encoded as three columns - 'MobilePhone', 'Internet' and 'Laptop'                                                                                                                                        |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Own a Mobile Phone                                                                                                                                                   | To understand the socioeconomic status of respondent                               | Encoded as two columns - 'MobileYes' and 'MobileNo'                                                                                                                                                      |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| If Business is registered with government                                                                                                                            | To understand business compliance                                                  | Encoded as two columns - 'BusinessRegistered_Yes' and 'BusinessRegistered_No'                                                                                                                            |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Respondent's main income source                                                                                                                                      | To understand the socioeconomic status of respondent                               | Encoded as binary columns for each income type, such as Formal_sector_salaried, Traders_agriculatural_proudcts, etc.                                                                                     |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Who they discuss their financial matters with                                                                                                                        | To understand social support networks                                              | Encoded as 15 columns - 'Advice_HouseholdMember' and 'Advice_Bank' and so on.                                                                                                                            |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| What they did when they could not meet spending needs                                                                                                                | To understand coping mechanisms                                                    | Encoded as 10 columns - 'UsedSavings' and 'BorrowedFromBank'                                                                                                                                             |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Whether respondent is comfortable with banks                                                                                                                         | To gauge trust in financial institutions                                           | Encoded as 'ComfortableWithBank_Yes' and 'ComfortableWithBank_No'                                                                                                                                        |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Financial Education                                                                                                                                                  | To understand the respondent's financial literacy                                  | Encoded as 'ReceivedFinancialEducation_Yes' and 'ReceivedFinancialEducation_No'                                                                                                                          |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Retirement plans                                                                                                                                                     | To understand future financial planning                                            | Encoded as 'Old_ChildrenWillTakeCareOfMe' and 'Old_Savings'                                                                                                                                              |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Confidence in financial institutions                                                                                                                                 | To understand perceptions of financial institutions                                | Encoded as "ConfidentWithBanks" and 'ConfidentWithPensionFund' and so on                                                                                                                                 |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| If they are paying back money they owe within a year                                                                                                                 | To understand debt management                                                      | Encoded as PayingBack_Yes and PayingBack_No.                                                                                                                                                             |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| If they have borrowed money from any financial institutions                                                                                                          | To understand access to credit                                                     | Encoded as binary columns for each financial institution type such as 'BorrowedFrom_Bank' and 'BorrowedFrom_MFI'                                                                                         |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| How often a respondent use the payment type for purchases                                                                                                            | To understand spending habits                                                      | Encoded as binary columns for each payment type such as Used_Cash, Used_ATM, Used_CreditCard and so on.                                                                                                  |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Assistance from Government or NGO and problems they face receiving assistance                                                                                        | To understand reliance on assistance and issues                                    | Encoded as binary columns such as 'Basic_TASAF_Cash_Transfer ' and 'Money_Didnt_Come'                                                                                                                    |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Reporting problems regarding financial matters                                                                                                                       | To understand barriers in reporting issues                                         | Encoded as binary columns such as 'Do_Not_Know_Where_To_Go' and 'Do_Not_Think_They_Care'                                                                                                                 |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Confidence and Trust in Insurance Companies                                                                                                                          | To assess trust in insurance                                                       | Encoded as relevant indicators such as Trust_Insurance_Companies                                                                                                                                         |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Insurance ownership and type of insurance                                                                                                                            | To understand financial protection                                                 | Encoded as binary columns for each insurance type such as Motor_vehicle_insurance, Health_Insurance, etc.                                                                                                |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| If they have any issues accessing financial products                                                                                                                 | To understand barriers to access                                                   | Network_Failure and Agent_Absent                                                                                                                                                                         |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Access and relationship with moneylender                                                                                                                             | To understand trust and access to informal credit                                  | Moneylender_Access and Moneylender_TreatsWell                                                                                                                                                            |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| If respondent is involved with a Bank, MFI, PENSION, INSURANCE, SACCO, CAPITAL Fund Managers, FORMAL INVESTMEN, CMG, INFORMAL MONEYLENDER or apart of a Social group | To assess engagement with financial institutions, microfinance, pension and so on. | Encoded as Banked and Not_Banked, MFI and Not_MFI and so on.                                                                                                                                             |
+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

## Code

```{r}
#| eval: False
tanzania_encoded <- tanzania_fin %>%
  mutate(districts = tanzania_fin$dist_name)%>%
  mutate(dist_name = as.factor(dist_name)) %>%
  mutate(value = 1) %>% 
  pivot_wider(names_from = dist_name, 
              values_from = value, 
              values_fill = 0, 
              names_prefix = "dist_") %>%
  mutate(
    Urban = ifelse(`Cluster.type` == "Urban", 1, 0),
    Rural = ifelse(`Cluster.type` == "Rural", 1, 0)) %>%
  mutate(
    HHH = ifelse(`C2..Who.is.the.head.of.this.household.` == "Respondent is hhh", 1, 0), 
    NotHHH = ifelse(`C2..Who.is.the.head.of.this.household.` == "Respondent not hhh", 1, 0))%>%
  mutate(
    Male = ifelse(`C9..Respondent.gender.Interviewer.to.observe` == "Male", 1, 0), 
    Female = ifelse(`C9..Respondent.gender.Interviewer.to.observe` == "Female", 1, 0))%>% 
  mutate(
    NoFormalEducation = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "No formal education", 1, 0),
    SomePrimary = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Some primary", 1, 0),
    PrimaryCompleted = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Primary completed", 1, 0),
    PostPrimaryTechnicalTraining = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Post primary technical training", 1, 0),
    SomeSecondary = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Some secondary", 1, 0),
    SecondaryCompleted_OLevel = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Secondary competed-O level", 1, 0),
    SomeUniversityOrHigherEducation = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "Some University or other higher education", 1, 0),
    UniversityOrHigherEducationCompleted = ifelse(`C11..Highest.level.of.education.completed..Can.read.out` == "University or higher education completed", 1, 0))%>%
  mutate(
    MobilePhone = ifelse(`C23..Which.of.the.following.do.you.have.access..to..Mobile.phone` == "Yes", 1, 0),
    Internet = ifelse(`C23..Which.of.the.following.do.you.have.access..to..Internet` == "Yes", 1, 0),
    Laptop = ifelse(`C23..Which.of.the.following.do.you.have.access..to..Laptop.or.desktop.or.tablet` == "Yes", 1, 0))%>%
  mutate(
    MobileYes = ifelse(`C24.1..Do.you.personally.own.a.mobile.phone.` == "Yes", 1, 0),
    MobileNo = ifelse(`C24.1..Do.you.personally.own.a.mobile.phone.` == "No", 1, 0)
  ) %>%
  mutate(
    BusinessRegistered_Yes = ifelse(`X6.4a.Is.your.business.registered.with.any.government.authority.` == "Yes", 1, 0),
    BusinessRegistered_No = ifelse(`X6.4a.Is.your.business.registered.with.any.government.authority.` == "No", 1, 0)
  ) %>%
  mutate(
    TriedButNot = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..Tried.but.was.not` == "Yes", 1, 0),
    NoTime = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..Don.t.have.time` == "Yes", 1, 0),
    NoMoney = ifelse(`X6.4c..Why.you.have.not.registered.this.business..Dont.have.money` == "Yes", 1, 0),
    TooComplicated = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..It.is.too.complicated` == "Yes", 1, 0),
    NoBenefit = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..No.benefit` == "Yes", 1, 0),
    BusinessTooSmall = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..Business.is.too.small` == "Yes", 1, 0),
    NoPayment = ifelse(`X6.4c..Why.you.have.not.registered.this.business..Don.t.want.to.pay` == "Yes", 1, 0),
    DontKnowHow = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..Dontt.know.how` == "Yes", 1, 0),
    RegistrationInProcess = ifelse(`X6.4c.Please.tell.me.why.you.have.not.registered.this.business..Registration.is.being.processed` == "Yes", 1, 0))%>%
  mutate(
    Formal_sector_salaried = ifelse(Main.Income.Source == "Formal sector salaried", 1, 0),
    Informal_sector_salaried = ifelse(Main.Income.Source == "Informal sector salaried", 1, 0),
    Farmers_and_fishers = ifelse(Main.Income.Source == "Farmers and fishers", 1, 0),
    Traders_agricultural_products = ifelse(Main.Income.Source == "Traders - agricultural products", 1, 0),
    Traders_non_agricultural = ifelse(Main.Income.Source == "Traders - non-agricultural", 1, 0),
    Service_providers = ifelse(Main.Income.Source == "Service providers", 1, 0),
    Piece_work_casual_labor = ifelse(Main.Income.Source == "Piece work/casual labor", 1, 0),
    Rental_income = ifelse(Main.Income.Source == "Rental income", 1, 0),
    Interest_from_savings_investments_stocks_unit_trusts = ifelse(Main.Income.Source == "Interest from savings, investments, stocks, unit trusts etc.", 1, 0),
    Pension = ifelse(Main.Income.Source == "Pension", 1, 0),
    Welfare = ifelse(Main.Income.Source == "Welfare", 1, 0),
    Gambling = ifelse(Main.Income.Source == "Gambling", 1, 0),
    Dependents = ifelse(Main.Income.Source == "Dependents", 1, 0),
    Other = ifelse(Main.Income.Source == "Other", 1, 0)
  )%>%
  mutate(
    Advice_HouseholdMember = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..A.household.member` == "Yes", 1, 0),
    Advice_FamilyMember = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Another.family.member` == "Yes", 1, 0),
    Advice_Employer = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Employer` == "Yes", 1, 0),
    Advice_Bank = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Bank` == "Yes", 1, 0),
    Advice_Microfinance = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Microfinance.institution` == "Yes", 1, 0),
    Advice_SACCOS = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Savings.and.credit.cooperative..SACCOS.` == "Yes", 1, 0),
    Advice_FinancialAdvisor = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Financial.advisor` == "Yes", 1, 0),
    Advice_FarmersAssociation = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Farmers.association` == "Yes", 1, 0),
    Advice_BusinessAssociation = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Business.association` == "Yes", 1, 0),
    Advice_SavingsGroup = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Savings.group` == "Yes", 1, 0),
    Advice_Moneylender = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Moneylender.in.community` == "Yes", 1, 0),
    Advice_GovernmentOfficial = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Government.official` == "Yes", 1, 0),
    Advice_VillageElder = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Village.elder.elder` == "Yes", 1, 0),
    Advice_OtherSpecify = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..Other.specify` == "Yes", 1, 0),
    Advice_NoOne = ifelse(`E.3.3..Who.do.you.talk.with.for.advice.about.money.matters..I.do.not.speak.with.anyone` == "Yes", 1, 0)
  )%>%
  mutate(
    UsedSavings = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Used.savings` == "Yes", 1, 0),
    BorrowedFromBank = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Borrowed.from.a.bank.MFI.SACCO` == "Yes", 1, 0),
    BorrowedThroughMM = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Borrowed.money.through.MM` == "Yes", 1, 0),
    BorrowedFromMoneyLender = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Borrowed.from.money.lender` == "Yes", 1, 0),
    BorrowedFromFriendFamily = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Borrowed.from.friend.family.community` == "Yes", 1, 0),
    GotAssistanceFromFamily = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Got.assistance.from.family.friends.community..that.I.did.not.need.to.repay.` == "Yes", 1, 0),
    SoldLivestock = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Sold.livestock` == "Yes", 1, 0),
    CutSpending = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Cut.spending` == "Yes", 1, 0),
    WorkedMore = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Worked.more` == "Yes", 1, 0),
    Others = ifelse(`E.7.N.1..Last.12.months...what.did.you.do.when.you.couldn.t.meet.your.regular.spending.needs._Others` == "Yes", 1, 0)
  )%>%
  mutate(
    ComfortableWithBank_Yes = ifelse(`E.5.1..You.feel.comfortable.to.go.into.a.bank.or.another.financial.institution` == "Yes", 1, 0),
    ComfortableWithBank_No = ifelse(`E.5.1..You.feel.comfortable.to.go.into.a.bank.or.another.financial.institution` == "No", 1, 0), ReceivedFinancialEducation_Yes = ifelse(`E.5.2..You.have.received.some.financial.education.in.your.life` == "Yes", 1, 0),
    ReceivedFinancialEducation_No = ifelse(`E.5.2..You.have.received.some.financial.education.in.your.life` == "No", 1, 0),
    Old_Savings = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Savings", 1, 0),
    Old_ChildrenWillTakeCareOfMe = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Children will take care of me", 1, 0),
    Old_MoneyFromFriendsRelatives = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Money from friends/relatives", 1, 0),
    Old_LandProperty = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Land/property", 1, 0),
    Old_OwnBusiness = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Own business", 1, 0),
    Old_RentalIncome = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Rental income", 1, 0),
    Old_DividendsFromShares = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Dividends from shares", 1, 0),
    Old_FarmingAgricultureLivestock = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Farming/agriculture/livestock", 1, 0),
    Old_Pension = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Pension", 1, 0),
    Old_InsurancePolicy = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Insurance policy", 1, 0),
    Old_NoPlans = ifelse(`E.10..How.will.you.mainly.ensure.that.you.have.money.to.meet.your.needs.when.you.are.old.and.can.no.longer.work.` == "Don’t know/Have no plans/Have not yet thought about how/where I would get the money for this purpose", 1, 0),
    ConfidentWithBanks = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._Banks` == "Yes", 1, 0),
    ConfidentWithMFIs = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._MFIs` == "Yes", 1, 0),
    ConfidentWithSACCOs = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._SACCOs` == "Yes", 1, 0),
    ConfidentWithMobileMoney = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._Mobile.money.operators` == "Yes", 1, 0),
    ConfidentWithPensionFund = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._Pension.fund` == "Yes", 1, 0),
    ConfidentWithSavingsGroups = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._Savings.groups.CMGs` == "Yes", 1, 0),
    ConfidentWithFamilyFriends = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._Family.friends` == "Yes", 1, 0),
    ConfidentWithUTTs = ifelse(`F.3.2.1..I.will.read.for.you.a.list.of.institutions..and.pease.tell.me.which.you.feel.confident.with.SAVINGS.your.money.with._UTTs` == "Yes", 1, 0),
    PayingBack_Yes = ifelse(`G.2.4..Have.you..in.the.past.12.months..been.paying.back.money.that.you.borrowed.` == "Yes", 1, 0),
    PayingBack_No = ifelse(`G.2.4..Have.you..in.the.past.12.months..been.paying.back.money.that.you.borrowed.` == "No", 1, 0),
    BorrowedFrom_Bank = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Bank` == "Yes", 1, 0),
    BorrowedFrom_MFI = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Microfinance.institution.such.as` == "Yes", 1, 0),
    BorrowedFrom_SACCO = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..SACCOS` == "Yes", 1, 0),
    BorrowedFrom_MobileMoney = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..A.mobile.money.service.provider` == "Yes", 1, 0),
    BorrowedFrom_PensionFund = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..A.pension.fund` == "Yes", 1, 0),
    BorrowedFrom_Employer = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Your.employer` == "Yes", 1, 0),
    BorrowedFrom_Family_PayBack = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Family.friends.that.you.had.to.pay.back` == "Yes", 1, 0),
    BorrowedFrom_Family_NoPayBack = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Family.friends.that.you.did.not.have.to.pay.back` == "Yes", 1, 0),
    BorrowedFrom_SavingsGroup = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Savings.group` == "Yes", 1, 0),
    BorrowedFrom_CommunityLender = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Someone.in.the.community.who.lend` == "Yes", 1, 0),
    BorrowedFrom_OtherSource = ifelse(`G.5.2.Did.you.borrow.money.from.the.following..Did.you.borrow.money.from.another.source` == "Yes", 1, 0),
    Used_Cash = ifelse(`X2.1.In.the.past.12.months..how.often.did.you.use.the.Cash.for.purchases.of.goods` == "Yes", 1, 0),
    Used_ATM_DebitCard = ifelse(`X2.2.In.the.past.12.months..how.often.did.you.use.the.ATM...Debit.card.for.purchases.of.goods` == "Yes", 1, 0),
    Used_CreditCard = ifelse(`X2.3.In.the.past.12.months..how.often.did.you.use.the.Credit.card.for.purchases.of.goods` == "Yes", 1, 0),
    Used_BankTransfer = ifelse(`X2.4.In.the.past.12.months..how.often.did.you.use.the.Bank.transfer.for.purchases.of.goods` == "Yes", 1, 0),
    Used_MobileMoney = ifelse(`X2.5.In.the.past.12.months..how.often.did.you.use.the.Mobile.money.for.purchases.of.goods` == "Yes", 1, 0),
    Used_Cryptocurrency = ifelse(`X2.6.In.the.past.12.months..how.often.did.you.use.the.Cryptocurrency.for.purchases.of.goods` == "Yes", 1, 0),
    Used_QR_Code = ifelse(`X3.1.In.the.past.12.months..how.often.did.you.use.the.QR.code..for.purchases.of.goods` == "Yes", 1, 0),
    Used_POS = ifelse(`X3.2.In.the.past.12.months..how.often.did.you.use.the.POS.for.purchases.of.goods` == "Yes", 1, 0),
    Used_LipaNamba = ifelse(`X3.3.In.the.past.12.months..how.often.did.you.use.the.Lipa.Namba.for.purchases.of.goods` == "Yes", 1, 0),
    Used_AZAMPay = ifelse(`X3.4.In.the.past.12.months..how.often.did.you.use.the.AZAM.Pay.for.purchases.of.goods` == "Yes", 1, 0),
    Basic_TASAF_Cash_Transfer = ifelse(`X1..You.indicated.earlier.that.you.get.assistance.from.Government.or.NGO..please.tell.me.which...Basic.TASAF.cash.transfer` == "Yes", 1, 0),
    Additional_Conditional_TASAF_Cash_Transfer = ifelse(`X1..You.indicated.earlier.that.you.get.assistance.from.Government.or.NGO..please.tell.me.which...Additional.Conditional.TASAF.cash.transfer` == "Yes", 1, 0),
    Pension_for_Elders = ifelse(`X1..You.indicated.earlier.that.you.get.assistance.from.Government.or.NGO..please.tell.me.which...Pension.for.elders..Znz.only.` == "Yes", 1, 0),
    Others_Assistance = ifelse(`X1..You.indicated.earlier.that.you.get.assistance.from.Government.or.NGO..please.tell.me.which...Others` == "Yes", 1, 0),
    
    Payment_Was_Late = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._Payment.was.late` == "Yes", 1, 0),
    Gave_Money_To_Official = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._You.had.to.give.money.to.an.official` == "Yes", 1, 0),
    Money_Didnt_Come = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._Money.didn.t.come.at.all` == "Yes", 1, 0),
    Received_Less_Money = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._You.received.less.money.than.should.have` == "Yes", 1, 0),
    Household_Not_On_Register = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._Your.household.not.on.the.register` == "Yes", 1, 0),
    Unable_To_Collection_Point = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._You.were.not.able.to.go.to.the.collection.point.and.you.didn.t.get.the.payment` == "Yes", 1, 0),
    No_Proof_Of_Identity = ifelse(`X2.Please.tell.me.which.of.the.following.problems.do.you.experience.with.these.payments._You.didn.t.have.proof.of.identity.and.didn.t.g.et.the.money` == "Yes", 1, 0),
    Do_Not_Know_Where_To_Go = ifelse(`J1.7.Why.don.t.you.report.problems..I.do.not.know.where.to.go` == "Yes", 1, 0),
    Places_Not_In_My_Area = ifelse(`J1.7.Why.don.t.you.report.problems..The.places.are.not.in.my.area` == "Yes", 1, 0),
    Do_Not_Think_They_Care = ifelse(`J1.7.Why.don.t.you.report.problems..I.do.not.think.they.care.about` == "Yes", 1, 0),
     Trust_Insurance_Companies = ifelse(`X1.3.Please.tell.me.which.of.the.following.you.is.true.for.you...You.trust.insurance.companies.to.pay.out.when.something.does.go.wrong` == "True", 1, 0),
    Insurance_Good_Idea = ifelse(`X1.3.Please.tell.me.which.of.the.following.you.is.true.for.you...Insurance.is.a.good.idea.for.everyone` == "True", 1, 0),
    Need_Insurance = ifelse(`X1.3.Please.tell.me.which.of.the.following.you.is.true.for.you...You.don.t.need.insurance.as.you.get.money.if.things.go.wrong` == "True", 1, 0),
    Have_Insurance_Cover = ifelse(`X2.Do.you.have.insurance.cover.` == "Yes", 1, 0),
    Motor_Vehicle_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Motor.vehicle.insurance` == "Yes", 1, 0),
    Households_Contents_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Households.contents.insura` == "Yes", 1, 0),
    Private_Health_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Private.health.insurance` == "Yes", 1, 0),
    National_Health_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..National.health.insurance` == "Yes", 1, 0),
    Community_Health_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Community.health.insurance` == "Yes", 1, 0),
    TIKA_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..TIKA..Tiba.Kwa.Kadi.` == "Yes", 1, 0),
    Social_Health_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Social.Health.Insurance.Be` == "Yes", 1, 0),
    Personal_Accident_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Personal.accident.insuranc` == "Yes", 1, 0),
    Life_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Life.insurance` == "Yes", 1, 0),
    Loan_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Loan.insurance.in.case.of` == "Yes", 1, 0),
    Building_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Building.insurance` == "Yes", 1, 0),
    Education_Policy = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Education.policy` == "Yes", 1, 0),
    Funeral_Cover = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Funeral.cover` == "Yes", 1, 0),
    Agricultural_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Agricultural.insurance` == "Yes", 1, 0),
    Ushirika_Afya = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Ushirika.afya` == "Yes", 1, 0),
    Motorcycle_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Motorcycle.insurance` == "Yes", 1, 0),
    Livestock_Insurance = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Livestock.insurance` == "Yes", 1, 0),
    Other_Insurance_Products = ifelse(`X3.1.1.Please.tell.me.which.of.the.following.you.have..Any.other.insurance.produc` == "Yes", 1, 0),
    Network_Failure = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Network.failure` == "Yes", 1, 0),
    Agent_Absent = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Agent.absent.not.available` == "Yes", 1, 0),
    Agent_No_Cash_Float = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Agent.didn.t.have.enough.cash.float` == "Yes", 1, 0),
    Agent_Overcharged = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Agent.overcharged.for.transactions` == "Yes", 1, 0),
    Lost_Money = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Lost.money` == "Yes", 1, 0),
    Agent_Requested_PIN = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Agent.wanted.you.to.reveal.your.PIN.number` == "Yes", 1, 0),
    Other_Problems = ifelse(`X10.1.In.the.past.12.months.have.you.experienced.any.of.the.following.when.you.wanted.to.use.mobile.money._Any.other.problem` == "Yes", 1, 0),
    Moneylender_Access = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..You.can.use.or.access.your.moneylender.whenever.you.need.to` == "Agree", 1, 0),
    Moneylender_TreatsWell = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..Your.moneylender.always.treats.you.well` == "Agree", 1, 0),
    Moneylender_InfoClear = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..Before.you.took.out.a.loan.with.your.moneylender..you.got.as.much.info.as.needed.and.it.was.easy.to.understand` == "Agree", 1, 0),
    Moneylender_NeedsMet = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..The.products.and.services.offered.by.your.moneylender.meets.your.financial.needs` == "Agree", 1, 0),
    Moneylender_FeesReasonable = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..Your.moneylender.s.fees.and.charges.are.reasonable` == "Agree", 1, 0),
    Moneylender_CustomerSupport = ifelse(`X4b.Please.tell.me.which.of.the.following.statements.you.agree.with..You.are.satisfied.with.your.moneylender.s.customer.support` == "Agree", 1, 0),
    Banked = ifelse(Banked == "Yes", 1, 0),
    Not_Banked = ifelse(Banked == "No", 1, 0),
    MFI = ifelse(MFI == "Yes", 1, 0),
    Not_MFI = ifelse(MFI == "No", 1, 0),
    PENSION = ifelse(PENSION == "Yes", 1, 0),
    Not_PENSION = ifelse(PENSION == "No", 1, 0),
    INSURANCE = ifelse(INSURANCE == "Yes", 1, 0),
    Not_INSURANCE = ifelse(INSURANCE == "No", 1, 0),
    SACCO = ifelse(SACCO == "Yes", 1, 0),
    Not_SACCO = ifelse(SACCO == "No", 1, 0),
    CAPITALM_FUND_MANAGERS = ifelse(CAPITALM_FUND_MANAGERS == "Yes", 1, 0),
    Not_CAPITALM_FUND_MANAGERS = ifelse(CAPITALM_FUND_MANAGERS == "No", 1, 0),
    FORM_INVESTMENTS = ifelse(FORM_INVESTMENTS == "Yes", 1, 0),
    Not_FORM_INVESTMENTS = ifelse(FORM_INVESTMENTS == "No", 1, 0),
    CMG = ifelse(CMG == "Yes", 1, 0),
    Not_CMG = ifelse(CMG == "No", 1, 0),
    INFORMAL_MONEYLENDER = ifelse(INFORMAL_MONEYLENDER == "Yes", 1, 0),
    Not_INFORMAL_MONEYLENDER = ifelse(INFORMAL_MONEYLENDER == "No", 1, 0),
    SOCIAL_GROUPS = ifelse(SOCIAL_GROUPS == "Yes", 1, 0),
    Not_SOCIAL_GROUPS = ifelse(SOCIAL_GROUPS == "No", 1, 0)
  )
```
:::

```{r}
#| eval: False
#| echo: False
write_rds(tanzania_encoded,"data/rds/tanzania_encoded")
```

```{r}
#| echo: False
tanzania_encoded <- read_rds("data/rds/tanzania_encoded")
```

Lastly, we will select the cleaned encoded columns as our final dataframe and replace all the missing values in encoded variables with 0. We will exclude geometry column in the process with the following code.

```{r}
#| eval: False
tanzania_cleaned <- tanzania_encoded %>%
  select(7,212:544)%>%
  mutate(across(-c(geometry), ~ ifelse(is.na(.), 0, .)))
```

There is now 13 records with age 0. We will impute age with the mean of the remaining respondents.

```{r}
#| eval: False
mean_age <- mean(tanzania_cleaned$Age, na.rm = TRUE)
tanzania_cleaned$Age[tanzania_cleaned$Age == 0] <- mean_age
```

Now, our dataframe is ready and cleaned with our chosen variables. We will save it.

::: panel-tabset
## Save

```{r}
#| eval: False
write_rds(tanzania_cleaned,"data/rds/tanzania_cleaned.rds")
```

## Load

```{r}
tanzania_cleaned <- read_rds("data/rds/tanzania_cleaned.rds")
```
:::

## 

# References

Demirgüç-Kunt, A., Klapper, L., & Singer, D. (2017). *Financial inclusion and inclusive growth: A review of recent empirical evidence* (WPS8040). Development Research Group, Finance and Private Sector Development Team. https://doi.org/10.1596/1813-9450-8040

Koomson, I., Churchill, S.A. & Munyanyi, M.E. Gambling and Financial Stress. *Soc Indic Res***163**, 473–503 (2022). https://doi.org/10.1007/s11205-022-02898-6

United Nations Capital Development Fund (UNCDF). (2024). *Insights from Tanzania: Global data reveals financial inclusion for women*. Retrieved from <https://policyaccelerator.uncdf.org/whats-new/insights-tanzania-global-data-reveals-financial-inclusion-women>

World Bank. (17 April, 2024). *Financial inclusion in Sub-Saharan Africa: Overview*. Retrieved from <https://www.worldbank.org/en/publication/globalfindex/brief/financial-inclusion-in-sub-saharan-africa-overview>

World Bank. (2022). *Global financial inclusion: A comprehensive overview*. Retrieved from <https://documents1.worldbank.org/curated/en/099305408252329233/pdf/SECBOS1e6ff9a00041b4261bcdc9c154a7b.pdf>
