---
title: "In Class Exercise 05"
author: "Sei Sar Hla Kyi"
date: "26 September 2024"
date-modified: "26 September 2024"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

Loading packages and Importing Data

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse)
```

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

Joining the two datasets

```{r}
hunan_GDPPC <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

Plotting choropleth map

```{r}
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC", 
          style = "quantile", 
          palette = "Blues",
          title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per capita by county, Hunan Province",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

## Global Measures of Spatial Association

### Step 1: Deriving Queen’s contiguity weights: sfdep methods

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1) 
```

```{r}
wm_q
```

### Computing Global Moran’ I

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
glimpse(moranI)
```

### Performing Global Moran’sI test

```{r}
global_moran_test(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
```

::: callout-tip
## The default for alternative argument is “two.sided”. Other supported arguments are “greater” or “less”. randomization, and By default the randomization argument is TRUE. If FALSE, under the assumption of normality.
:::

### Performing Global Moran’I permutation test

```{r}
set.seed(1234)
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

The statistical report on previous tab shows that the p-value is smaller than alpha value of 0.05. Hence, we have enough statistical evidence to reject the null hypothesis that the spatial distribution of GPD per capita are resemble random distribution (i.e. independent from spatial). Because the Moran’s I statistics is greater than 0. We can infer that the spatial distribution shows sign of clustering.

::: callout-important
## The numbers of simulation is alway equal to nsim + 1. This mean in nsim = 99. This mean 100 simulation will be performed.
:::

## LISA map

LISA map is a categorical map showing **outliers** and **clusters**. There are two types of outliers namely: High-Low and Low-High outliers. Likewise, there are two type of clusters namely: High-High and Low-Low cluaters. In fact, LISA map is an interpreted map by combining local Moran’s I of geographical areas and their respective p-values.

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

### Visualising local Moran’s I

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(
    main.title = "local Moran's I of GDPPC",
    main.title.size = 2)
```

### Visualising p-value of local Moran’s I

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_ii_sim") + 
  tm_borders(alpha = 0.5) +
   tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 2)
```

::: callout-warning
## For p-values, the appropriate classification should be 0.001, 0.01, 0.05 and not significant instead of using default classification scheme.
:::

### Visualising local Moran’s I and p-value

```{r}
tmap_mode("plot")
map1 <- tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(lisa) +
  tm_fill("p_ii",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### Plotting LISA map

In lisa sf data.frame, we can find three fields contain the LISA categories. They are *mean*, *median* and *pysal*. In general, classification in *mean* will be used as shown in the code chunk below.

```{r}
lisa_sig <- lisa  %>%
  filter(p_ii_sim < 0.05)
tmap_mode("plot")
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

## Hot Spot and Cold Spot Area Analysis (HCSA)

HCSA uses spatial weights to identify locations of statistically significant hot spots and cold spots in an spatially weighted attribute that are in proximity to one another based on a calculated distance. The analysis groups features when similar high (hot) or low (cold) values are found in a cluster. The polygon features usually represent administration boundaries or a custom grid structure.

### Computing local Gi\* statistics

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```

-   ::: callout-note
    ## Gi\* and local Gi\* are distance-based spatial statistics. Hence, distance methods instead of contiguity methods should be used to derive the spatial weight matrix. Since we are going to compute Gi\* statistics, include_self()is used.
    :::

### Computing local Gi\* statistics

```{r}
HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA
```

### Visualising Gi\*

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

### Visualising p-value of HCSA

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("p_sim") + 
  tm_borders(alpha = 0.5)
```

### Visuaising local HCSA

```{r}
tmap_mode("plot")
map1 <- tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HCSA) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### Visualising hot spot and cold spot areas

Now, we are ready to plot the significant (i.e. p-values less than 0.05) hot spot and cold spot areas by using appropriate tmap functions as shown below.

```{r}
HCSA_sig <- HCSA  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_sig) +
  tm_fill("cluster") + 
  tm_borders(alpha = 0.4)
```

::: callout-note
## Note

Figure above reveals that there is one hot spot area and two cold spot areas. Interestingly, the hot spot areas coincide with the High-high cluster identifies by using local Moran’s I method in the earlier sub-section.
:::
