---
title: "In-class Exercise 1"
author: "Sei Sar Hla Kyi"
date: "31 August 2024"
date-modified: "31 August 2024"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

Install and load packages

```{r}
pacman::p_load(tidyverse, sf, tmap, ggstatsplot)
```

Import data

```{r}
mpsz14_shp <- st_read(dsn = "data/",
                layer = "MP14_SUBZONE_WEB_PL")

```

```{r}
st_write(mpsz14_shp, 
         "data/MP14_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE)
```

```{r}
preschool_kml <- st_read("data/PreSchoolsLocation.kml")
preschool_geojson <- st_read("data/PreSchoolsLocation.geojson") 
```

Importing Shapefile

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                layer = "MPSZ-2019")
```

```{r}
mpsz19_kml <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

check the project of the imported sf objects.

```{r}
st_crs(mpsz19_shp)
```

import the Master Plan Sub-zone 2019 and Pre-schools Location with proper transformation

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
preschool <- st_read("data/PreSchoolsLocation.kml") %>%
  st_transform(crs = 3414)
```

Count the number of pre-schools in each planning sub-zone.

```{r}
mpsz19_shp <- mpsz19_shp %>%
  mutate(`PreSch Count` = lengths(
    st_intersects(mpsz19_shp, preschool)))
```

-   Derive the area of each planning sub-zone.

-   Drop the unit of measurement of the area (i.e. m\^2)

-   Calculate the density of pre-school at the planning sub-zone level.

    ```{r}
    mpsz19_shp <- mpsz19_shp %>%
      mutate(Area = units::drop_units(
        st_area(.)),
        `PreSch Density` = `PreSch Count` / Area * 1000000
      )
    ```

Exploratory Data Analysis (EDA) and Confirmatory Data Analysis (CDA)

```{r}
mpsz19_shp$`PreSch Density` <- as.numeric(as.character(mpsz19_shp$`PreSch Density`))
mpsz19_shp$`PreSch Count` <- as.numeric(as.character(mpsz19_shp$`PreSch Count`)) 
mpsz19_shp_1 <- as.data.frame(mpsz19_shp)

ggscatterstats(data = mpsz19_shp_1,
               x = `PreSch Density`,
               y = `PreSch Count`,
               type = "parametric",
               label.var = "SUBZONE_N",
               label.expression = `PreSch Density` > 40 & `PreSch Count` > 70)

```

```{r}
popdata <- read_csv("data/respopagesextod2023.csv")
```

```{r}
popdata2023 <- popdata %>% 
  group_by(PA, SZ, AG) %>% 
  summarise(`POP`=sum(`Pop`)) %>%  
  ungroup() %>% 
  pivot_wider(names_from=AG,
              values_from = POP)

colnames(popdata2023)
```

derive a tibble data.framewith the following fields PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY where by:

YOUNG: age group 0 to 4 until age groyup 20 to 24, ECONOMY ACTIVE: age group 25-29 until age group 60-64, AGED: age group 65 and above, TOTAL: all age group, and DEPENDENCY: the ratio between young and aged against economy active group.

```{r}
popdata2023 <- popdata2023 %>%
  mutate(YOUNG=rowSums(.[3:6]) # Aged 0 - 24, 10 - 24
         +rowSums(.[14])) %>% # Aged 5 - 9
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+ # Aged 25 - 59
  rowSums(.[15])) %>%  # Aged 60 -64
  mutate(`AGED`=rowSums(.[16:21])) %>%
  mutate(`TOTAL`=rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY`=(`YOUNG` + `AGED`)
  / `ECONOMY ACTIVE`) %>% 
  select(`PA`, `SZ`, `YOUNG`, 
         `ECONOMY ACTIVE`, `AGED`,
         `TOTAL`, `DEPENDENCY`)
```

Joining popdata2023 and mpsz19_shp

```{r}
popdata2023 <- popdata2023 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) 
```

```{r}
mpsz_pop2023 <- left_join(mpsz19_shp, popdata2023,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
pop2023_mpsz <- left_join(popdata2023, mpsz19_shp, 
                          by = c("SZ" = "SUBZONE_N"))
```

## **Analytical Map: Percentile Map**

```{r}
mpsz_pop2023 <- mpsz_pop2023 %>%
  drop_na()
```

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```


```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(mpsz_pop2023) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```


```{r}
percentmap("DEPENDENCY", mpsz_pop2023)
```

## **Analytical Map: Box Map**

```{r}
ggplot(data = mpsz_pop2023,
       aes(x = "",
           y = DEPENDENCY)) +
  geom_boxplot()
```

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
```

```{r}
boxmap("DEPENDENCY", mpsz_pop2023)
```

Plotting Interactive Map

```{r}
tmap_options(check.and.fix = TRUE)
tmap_mode("view")
boxmap("DEPENDENCY", mpsz_pop2023)
```
