---
title: "In Class Exericse 09"
author: "Sei Sar Hla Kyi"
date: "`r Sys.Date()`"
date-modified: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# Load Packages

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

# Import Data Packages

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL") %>%
  st_transform(crs = 3414)

hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") %>%
  st_transform(crs = 3414)

eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") %>%
  st_transform(crs = 3414)
```

## OD Matrix

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

# Data Cleaning and Updating Attributes

## Supply

```{r}
eldercare <- eldercare %>%
  select (fid,ADDRESSPOS) %>%
  mutate (capacity =100)
```

Set to 100 first to calibrate he model.

## Demand

```{r}
hexagons <- hexagons %>%
  select(fid)%>%
  mutate(demand = 100)
```

## OD Matrix

```{r}
distmat <- ODMatrix %>%
  select (origin_id,destination_id,total_cost)%>%
  spread(destination_id, total_cost) %>%
  select (-c(origin_id))
```

```{r}
distmat_km <- as.matrix(distmat/1000)
```

# Handsen's Accessibility

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

```{r}
colnames(acc_Hansen) <- "accHansen"
```

Next, we will convert the data table into tibble format by using the code chunk below.

```{r}
acc_Hansen <- tbl_df(acc_Hansen)
```

Lastly, *bind_cols()* of dplyr will be used to join the acc_Hansen tibble data frame with the hexagons simple feature data frame. The output is called `hexagon_Hansen`.

```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

### **Visualising Hansen’s accessibility**

#### Extracting map extend

```{r}
mapex <- st_bbox(hexagons)
```

Boundary box to reserve the dimension.

```{r}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## **Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method**

```{r}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- tbl_df(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

```{r}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 1,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

```{r}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

```{r}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```
